---
title: "Case Study 1 Thundercats"
output: html_notebook
---
```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE)
```
# Beer: 
   Humans have been consuming alcohol in its various forms for millenia. Fermentation
   is an art that we have perfected as a species. Beer ddrinkers are notoriously picky about
   the type of beer they consume, many prefer lower alcohol content varieties becuase they 
   prefer to drink for longer periods of time, others prefer a stronger brew and prefer to 
   drink beer for its flavor content rather than its intoxicating effects. Other drinkers
   prefer both - a beer that is easy to drink (low ABV) but with more flavor/body (higher      IBU). 
   
### In which state can we find the strongest brew?

   Strength of beer is usually characterized by the ABV (Alcohol Content By Volume)
   
   By analyzing the distrubution of ABV by state, we will attempt to answer this important QOI
   
### Where is the beer most bitter?
   
   Bitterness of beer is usually characterized by the IBU (International Bitterness Unit)
   
   By analyzing the distrubution of IBU by state, we will attempt to answer this important QOI 
   
### Data
   
   To investiage these QOIs - we will use an observational study conducted across all 50 states (except North Dakota?) to rank 2305 indivdual brews on their strength and bitterness. Unfortunately - our dataset is rather sparsely populated for IBU data making any inferences on this variable subject to skepticism and further investigation.


```{r ECHO=TRUE}
library(sqldf)
library(ggplot2)
```

```{r}

beers = read.csv('Data/Beers.csv')
breweries = read.csv('Data/Breweries.csv')
states = read.csv('Data/StatesByRegion.csv')
breweries = sqldf('SELECT * FROM breweries b LEFT JOIN states s ON s.Code = b.State')
```

### Merge the beer and brew data

```{r}
merged_data = sqldf('SELECT *
                        FROM beers b
                        LEFT JOIN breweries br ON b.Brewery_id = br.Brew_ID

')
```

## Top 6 rows

```{r}
top_6 = data.frame(head(merged_data, 6))
top_6

```

## last 6 rows

```{r}
bottom_6 = data.frame(tail(merged_data, 6))
bottom_6
```

### Breweries by State
    
    Breweries seem to be concentrated highly in the top 5 states (CO, CA, MI, OR and TX)
    Comprising 175 of the 558 breweries in the study
    
```{r}
breweries_by_state = group_by(breweries, State) %>% summarise(
  count = n()
)
breweries_by_state[order(-breweries_by_state$count), ]
```

### By region?
    
    Micro brews abound out west!
    
    188 of the breweries in the study were located out west. If one were to target places to     sell beer, the left coast would be an ideal place to start
```{r}
breweries_by_region = group_by(breweries, Region) %>% summarise(
  count = n()
)
breweries_by_region[order(-breweries_by_region$count), ]

```
## Best Cities for a Beer Tour
  
  Want to check out a tour of local microbreweries? These cities are the top 5 when it 
  comes to craft beer specials!
  
  Portland here we come!
  
```{r}
breweries_by_city = group_by(breweries, City)  %>% summarise(
  count = n()
)

top_5_cities = data.frame(head(breweries_by_city[order(-breweries_by_city$count), ], 5))
top_5_cities
```



## NAs by column
  As noted in the introduction - IBU data is not widely available (1005 NA values)
  
  ABV - on the other hand - is required to be placed on the label of nearly all brews, hence   the data is much more reliable for that metric.
  
  Overall- the most represented (and potentially most popular) style of beer is the American   IPA (making up roughly 20% of the overall study)
  
  The state with the most beers sampled overall was Colorado. A state known for its liberal 
  policies surrounding consumption of quasi-illicit substances.
  
```{r}
summary(merged_data)
```

## effect of all the missing IBU data?
    
    Lots of beers were not categorized by IBU (almost half!)
    
    A large chunk of these beers were IPAs, which are generally assumed to have higher IBU values

```{r}
no_ibu = merged_data[is.na(merged_data$IBU), ]
summary(no_ibu)
```
## Average IBU by Beer Type

```{r}
beers_by_type = group_by(merged_data, Style)  %>% summarise(
  average_ibu = mean(IBU, na.rm = TRUE)
)
beers_by_type[order(-beers_by_type$average_ibu), ]
```


## summary stats by state
    
    Like strong flavored beer? Visit Maine! 
    
    Prefer to get a bit more buzzed? Check out the state capitol and have a beer with the       President! He's not always drunk - it just seems that way!
```{r}
medians_by_state = group_by(merged_data, State) %>% summarise(
  median_abv = median(ABV, na.rm=TRUE),
  median_ibu = median(IBU, na.rm=TRUE)
)

medians_by_region = group_by(merged_data, Region) %>% summarise(
  median_abv = median(ABV, na.rm=TRUE),
  median_ibu = median(IBU, na.rm=TRUE)
)

medians_by_state[order(-medians_by_state$median_ibu), ]

medians_by_state[order(-medians_by_state$median_abv), ]

medians_by_region[order(-medians_by_region$median_ibu), ]

medians_by_region[order(-medians_by_region$median_abv), ]
```

## Plot of Median ABV By State and Region
  
    Don't go to South Dakota if you like bitter beer, there may not be any....
    
    If you prefer a nice mix of ABV/IBU - head west, friend, there's gold out there!
```{r}

state_plot_abv = ggplot(medians_by_state, aes(x=reorder(medians_by_state$State, median_abv), y=medians_by_state$median_abv, fill=medians_by_state$State), palette=palette(rainbow(length(levels(medians_by_state$State))))) + geom_bar(stat = "identity") + coord_flip() + labs(x = 'State', y='Median ABV', fill='State') 

print(state_plot_abv + ggtitle('Median ABV By State') +
  theme(plot.title = element_text(hjust = 0.5)))


state_plot_ibu = ggplot(medians_by_state, aes(x=reorder(medians_by_state$State, median_ibu), y=medians_by_state$median_ibu, fill=medians_by_state$State), palette=palette(rainbow(length(levels(medians_by_state$State))))) + geom_bar(stat = "identity") + coord_flip() + labs(x = 'State', y='Median IBU', fill='State') 
print(state_plot_ibu + ggtitle('Median IBU By State') +
  theme(plot.title = element_text(hjust = 0.5)))


region_plot_abv = ggplot(medians_by_region, aes(x=reorder(medians_by_region$Region, median_abv), y=medians_by_region$median_abv, fill=medians_by_region$Region), palette=palette(rainbow(length(levels(medians_by_region$Region))))) + geom_bar(stat = "identity") + coord_flip() + labs(x = 'Region', y='Median ABV', fill='Region') 

print(region_plot_abv + ggtitle('Median ABV By Region') +
  theme(plot.title = element_text(hjust = 0.5)))


region_plot_ibu = ggplot(medians_by_region, aes(x=reorder(medians_by_region$Region, median_ibu), y=medians_by_region$median_ibu, fill=medians_by_region$Region), palette=palette(rainbow(length(levels(medians_by_region$Region))))) + geom_bar(stat = "identity") + coord_flip() + labs(x = 'Region', y='Median IBU', fill='Region') 
print(region_plot_ibu + ggtitle('Median IBU By Region') +
  theme(plot.title = element_text(hjust = 0.5)))

```

## most boozy beer state
  
  Colorado - no surprise here. Colorado is known for its liberal policies surrounding consumption of alcohol and other substances - hence having a beer with the strength of the average glass of wine is not altogether unexpected.
  
  Although - if you just want to go to a state and get drunk off of a variety of beers Kentucky may be your best bet with a median ABV > 0.06
  
```{r}
most_boozy = sqldf('SELECT * FROM merged_data ORDER BY ABV DESC')
most_boozy = head(most_boozy, 1)
most_boozy$State
most_boozy$Name
most_boozy$ABV
```

## most bitter beer state

   Oregon has the most bitter beer in the sample - the aptly named "Bitter Bitch Imperial IPA" clocking in at a whopping 138 IBU!
   
   BUT the most bitter beer on average in the sample (based off of the median) would be Maine

```{r}
most_bitter = sqldf('SELECT * FROM merged_data ORDER BY IBU DESC')
most_bitter = head(most_bitter, 1)
most_bitter$State
most_bitter$Name
most_bitter$IBU
```

## ABV Summary

    Visual inspection of the ABV data shows a strong positive skew towards the upper end of the range with quite a few outlyers.
    
```{r}
abv_summary = summary(merged_data$ABV)
abv_summary
ggboxplot(merged_data$ABV)
```

## scatter ABV/IBU Data

      SCatter/Regression shows a strong indication of a linear trend between ABV/IBU with higher values for IBU generally correlating to higher ABV content. Since so much of the IBU data is missing - we should impute the values for the missing IBU data with it's average by style

```{r}
ggscatter(merged_data, x='IBU', y='ABV', color = 'black'
          , add=c('reg.line')
          , conf.int = TRUE
          , title = 'ABV versus IBU')

```

## Lets impute some values and re-run!

    Imputing the average value by style allowed us to better analyze the trend between ABV/IBU in the absence of missing data. Most of the data that was missing is related to Ciders which should be excluded anyway since IBU is not a measure typically associated with that style of beverage.
```{r}
imputed_ibu = sqldf('

SELECT m.Name, m.ABV, COALESCE(m.IBU, b.average_ibu) aS Imp_IBU FROM merged_data m LEFT JOIN beers_by_type b ON b.Style = m.Style'
  
  
  
)
ggscatter(imputed_ibu, x='Imp_IBU', y='ABV', color = 'black'
          , add=c('reg.line')
          , conf.int = TRUE
          , title = 'ABV versus IBU (imputed)')
```